---
title: "simpleRegress"
author: "Alexis Velasco"
date: "6/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```



## now predict the length of a bear who has a neck of 17, 22 and 26

```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```

## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

Comparing length to weight to find the best prediction of length gave us a smaller r-squared than neck only giving us a r-squared value of 0.7472. Using weight would not be better than using neck to predict length.  

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

Using head length to predict length is better than using neck. It gave an adjusted r-squared value of 0.8434 a bigger value than when using neck. Also the residual standard error was smaller than neck with 5.376 head length's standard error was only 4.235. 

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

Using age to compare for length was worse than using neck. Having an adjusted r-squared value of 0.5166 but having a standard error of 7.511.the plot demonstrated this grately by seeing many points fall under the line.

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

Using head width to compare for length was not better than neck. Having a smaller adjusted r-squared of 0.5321 and a standard error of 7.32. The plot being so spread out between shows the effects of such a larger standard error.


```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

Using chest to compare for length prediction is better than using neck. With an adjusted r-squared value of 0.7869 and smaller standard error of 4.939. Chest would be better then using neck to predict the length of the bears.

```{r}
simp <- lm(LENGTH~MONTH,data=bears)
plot(LENGTH~MONTH,data=bears)
abline(simp)
summary.lm(simp)
```
 Using month to compare for length prediction would be way worse than using neck. With an adjusted r-squared value of -0.01153 and the standard error of 10.76. The plot really shows this by the points being so spread out.




Going through the different variables to see which gives the best prediction of bear's length , the results showed that using head length was the most efficient. With an adjusted r-squared of 0.8434 and only having a standard error of 4.235. The second best way to predict length would be using chest to compare, with a adjusted r-squared value of 0.7869 and a standard error of 4.939. Using head length and chest are both better than using neck to predict the length of these bears. Neck having standard error of 5.376 and an adjusted r-squared of 0.7524. Both head length and chest had a smaller standard error and a bigger adjusted r-squared. 
